---
title: "Diet.beta.div"
author: "Keri Kemp"
date: '2022-07-14'
output:
  pdf_document: default
  html_document: default
---
```{r, global_options, include=FALSE}
knitr::opts_chunk$set(message=FALSE, warning=FALSE, tidy.opts=list(width.cutoff=60)) 
knitr::opts_knit$set(root.dir = "~/Library/CloudStorage/Box-Box/_Mrug/Diet/R_analysis/Beta_diversity") #desktop
#knitr::opts_knit$set(root.dir = "~/Box/_Mrug/Diet/R_analysis") #laptop
```

```{r message=FALSE, include=FALSE}
#Load libraries
library(tidyverse)
library(phyloseq)
library(vegan)
library(ggplot2)
library(ggfortify)
library(ape)
```


### Import cleaned microbiome data into phyloseq
```{r message=FALSE, include=TRUE}

# read data in 
otu <- read.table("~/Library/CloudStorage/Box-Box/_Mrug/Diet/R_analysis/_bin/Diet_asv_table.txt",header=TRUE, row.names=1)
samples<-read.table("~/Library/CloudStorage/Box-Box/_Mrug/Diet/R_analysis/_bin/Diet_full_map.txt",sep="\t",header=T,row.names=2)

OTU = otu_table(otu, taxa_are_rows=FALSE)
Diet <- phyloseq(OTU,
               sample_data(samples)
               )

# Inspect the phyloseq object
Diet
# phyloseq-class experiment-level object
# otu_table()   OTU Table:         [ 1253 taxa and 137 samples ]
# sample_data() Sample Data:       [ 137 samples by 51 sample variables ]

```

### Calculate Bray-curtis Distance Matrix on rarefied ASV count data
```{r message=FALSE, include=TRUE}

set.seed(5455)
rare_ps <- phyloseq::rarefy_even_depth(Diet)
otu_rare = as.data.frame(as(otu_table(rare_ps), "matrix"))

bray_dist <- vegan::vegdist(otu_rare, method="bray", transformation = "hellinger")
jac_dist <- vegan::vegdist(otu_rare, method="jaccard", transformation = "hellinger")

```

### Permutational analysis of variance (PERMANOVA) using vegan's adonis2 function
The adonis function works by first finding the centroids for each group and then calculating the squared deviations of each of site to that centroid. Then significance tests are performed using F-tests based on sequential sums of squares from permutations of the raw data.

### I. PERMANOVA - Covariates

```{r}
adonis2(bray_dist~female, samples, permutations=999)

adonis2(bray_dist~c1zbmi, samples, permutations=999)

adonis2(bray_dist~FISMean, samples, permutations=999)

adonis2(bray_dist~SES, samples, permutations=999)
